An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-08-25 latex_files/ch3/3_2_2_training.tex
changed notation to be inline with ROMC paper

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimizer or Bayesian optimization. These tasks are
compleed sequentially, as shown in figure~\ref{fig:romc_overview}. The
definition of the optimisation problems is performed by drawing
\code{n_1} integer numbers from a discrete uniform distribution
$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
used in \pkg{ELFI}'s random simulator. The seed $u_i$ is the
computational analogous of the random variables $\vb_i$, that were
defined in the previous chapter; freezing the \code{seed} absorbs all
the randomness of the simulator transforming it to a deterministic
function. Setting the argument \code{use_bo=True}, chooses the
Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
case, apart from obtaining the optimal points $\thetab_i^*$, we also
fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
scenario, in the following steps, $\hat{d}_i(\thetab)$ will replace
$g_i(\thetab)$ as the distance function.

@@

\noindent
This function helps the user decide which threshold $\epsilon$ to
use. It plots a histogram of the distances at the optimal point
$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $d_i^*$ in case
\code{use_bo=True}. The function accepts all keyword arguments and
forwards them to the underlying \code{matplotlib.hist()} function; in
this way the user may customise some properties of the histogram, such
as the number of bins or the range of values.

@@

\noindent
This method constructs a bounding box around the optimal point
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian $\hess_i = \jac_i^T \jac_i$. The
eigenvectores are computed using the function
\code{numpy.linalg.eig()}. A check is performed so that the matrix
$\hess_i$ is not singular; if this is the case, the eigenvectors are
set to be the vectors of the standard Euclidean basis i.e.\
$\{ \mathbf{e_1} = (1, 0, \ldots), \mathbf{e_2} = (0,1,0,\ldots),
\text{etc} \}$. The limits of the bounding box are obtained by
repeteadly querying the distance function ($g_i(\thetab)$ or
$\hat{d}(\thetab)$) along the search directions given by $\hess_i$.

>>>>

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimizer or Bayesian optimization. These tasks are
compleed sequentially, as shown in figure~\ref{fig:romc_overview}. The
definition of the optimisation problems is performed by drawing
\code{n1} integer numbers from a discrete uniform distribution
$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
used in \pkg{ELFI}'s random simulator. The seed $u_i$ is the
computational analogous of the random variables $\vb_i$, that were
defined in the previous chapter; freezing the \code{seed} absorbs all
the randomness of the simulator transforming it to a deterministic
function. Setting the argument \code{use_bo=True}, chooses the
Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
case, apart from obtaining the optimal points $\thetab_i^*$, we also
fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
scenario, in the following steps, $\hat{d}_i(\thetab)$ replaces
$d_i(\thetab)$ as the distance function.

@@

\noindent
This function helps the user decide which threshold $\epsilon$ to
use. It plots a histogram of the distances at the optimal point
$d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $\hat{d}_i^*$ in case
\code{use_bo=True}. The function accepts all keyword arguments and
forwards them to the underlying \code{matplotlib.hist()} function; in
this way the user may customise some properties of the histogram, such
as the number of bins or the range of values.

@@

\noindent
This method constructs a bounding box around the optimal point
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian
$\hess_i = \jac_i^T \jac_i$\footnote{The Jacobian matrix is computed
  on the output of the summary statistics
  $\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))$ at $\thetab_i^*$} if
gradients have been used, or as the Hessian of the GP, in case of
Bayesian Optimisation. The eigenvectors, which are the axes of the
bounding box, are computed using the function
\code{numpy.linalg.eig()}. Afterwards, the limits along the
eigenvectors are computed with
Algorithm~\ref{alg:region_construction}.

### 2021-08-25 latex_files/graphs/romc_graph.tex
applied fixes
fixed warnings
changed notation to be inline with ROMC paper

\begin{figure}[h]
  \begin{center}
  %   \resizebox{.32\textwidth}{!}{
  %     \begin{tikzpicture}[node distance=1.4cm, scale=.1]
  %     \end{tikzpicture}
  %   }
    \resizebox{.8\textwidth}{!}{    
      \begin{tikzpicture}[node distance=1.2cm, scale=.1]
        % first graph
        % private functions
        \node (n1) [train_process, xshift=-.8cm] { $\_sample\_nuisance()$  };
        \node (n2) [train_process, below of=n1] { $\_define\_objectives()$  };
        \node (n3) [decision, below of=n2, yshift=-.5cm] { $grads$  };
        \node (n4) [train_process, left of=n3, yshift=-1cm, xshift=-1.5cm] { $\_solve\_gradients()$  };
        \node (n5) [train_process, right of=n3, yshift=-1cm, xshift=1.5cm] { $\_solve\_bo()$  };
        \node (n6) [train_process, below of=n3, yshift=-1.5cm] { $\_filter\_solutions()$  };
        \node (n7) [train_process, below of=n6] { $\_build\_boxes()$  };
        \node (n8) [decision, below of=n7] { $h?$  };
        \node (n9) [train_process, right of=n8, yshift=-.5cm, xshift=1cm] { $\_fit\_models()$  };

@@

        % second graph
        \node (pro1) [train_process, left of=n1, xshift=-7cm, yshift=-1cm, minimum width=4cm, minimum height=1cm] { define $d_i(\thetab) \forall i$  };
        \node (pro2) [train_process, below of=pro1, yshift=-1cm, minimum width=4cm, minimum height=1cm] { Solve $\thetab_i^*, d_i^*, \forall i$  };
        \node (filter) [train_process, below of=pro2, yshift=-1.5cm, minimum width=4cm, minimum height=1cm] { Filter solutions  };
        \node (proposal_region) [train_process, below of=filter, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $q_i \forall i$};
        \node (surrogate) [train_process, below of=proposal_region, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $\Tilde{d}_i \forall i$};
        \node (posterior) [train_process, below of=surrogate, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Define $p_{d,\epsilon}(\thetab)$};

@@

      \end{tikzpicture}
    }
  \end{center}
  \caption{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
\end{figure}

>>>>

\begin{figure}[ht]
  \begin{center}
  %   \resizebox{.32\textwidth}{!}{
  %     \begin{tikzpicture}[node distance=1.4cm, scale=.1]
  %     \end{tikzpicture}
  %   }
    \resizebox{.8\textwidth}{!}{    
      \begin{tikzpicture}[node distance=1.2cm, scale=.1]
        % first graph
        % private functions
        \node (n1) [train_process, xshift=-.8cm] { $\_sample\_nuisance()$  };
        \node (n2) [train_process, below of=n1] { $\_define\_objectives()$  };
        \node (n3) [decision, below of=n2, yshift=-.5cm] { $grads?$  };
        \node (n4) [train_process, left of=n3, yshift=-1cm, xshift=-1.5cm] { $\_solve\_gradients()$  };
        \node (n5) [train_process, right of=n3, yshift=-1cm, xshift=1.5cm] { $\_solve\_bo()$  };
        \node (n6) [train_process, below of=n3, yshift=-1.5cm] { $\_filter\_solutions()$  };
        \node (n7) [train_process, below of=n6] { $\_build\_boxes()$  };
        \node (n8) [decision, below of=n7] { $fit?$  };
        \node (n9) [train_process, right of=n8, yshift=-1cm, xshift=1cm] { $\_fit\_models()$  };

@@

        % second graph
        \node (pro1) [train_process, left of=n1, xshift=-7cm, yshift=-1cm, minimum width=4cm, minimum height=1cm] { define $d_i(\thetab) \forall i$  };
        \node (pro2) [train_process, below of=pro1, yshift=-1cm, minimum width=4cm, minimum height=1cm] { Solve $\thetab_i^*, d_i^*, \forall i$  };
        \node (filter) [train_process, below of=pro2, yshift=-1.5cm, minimum width=4cm, minimum height=1cm] { Filter solutions  };
        \node (proposal_region) [train_process, below of=filter, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $q_i \forall i$};
        \node (surrogate) [train_process, below of=proposal_region, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Fit $\Tilde{d}_i \forall i$};
        \node (posterior) [train_process, below of=surrogate, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Define $p_{d,\epsilon|\data}(\thetab)$};

@@

      \end{tikzpicture}
    }
  \end{center}
  \caption{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
\label{fig:romc_overview}
\end{figure}

### 2021-08-25 paper.tex
added argmin and argmax macros
plots after changes
changed notation to be inline with ROMC paper

@@

In this section, we present our implementation. At~\ref{subsec:ELFI}
we introduce the reader to the \proglang{Python} package \pkg{Engine
  for likelihood-free inference (ELFI)}
package. At~\ref{subsec:general_design} we explain the general design
that we follow. Then, we divide the implementation into three parts;
at~\ref{subsec:training} we present the training (fitting) part,
at~\ref{subsec:inference} the inference part and
at~\ref{subsec:evaluation} the evaluation part. Finally,
at~\ref{subsec:extensibility} we describe how a user may extend ROMC
with its custom modules.

@@

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}
\subsection{Moving average example}
\input{latex_files/ch4/4_2_ma2.tex}

\subsection{Execution time}
\input{latex_files/ch4/4_3_exec.tex}

>>>>

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

@@

In this section, we present the implementation. At~\ref{subsec:ELFI}
we introduce the reader to the \proglang{Python} package \pkg{Engine
  for likelihood-free inference
  (ELFI)}. At~\ref{subsec:general_design} we explain the general
design that we follow. Then, we divide the implementation into three
parts; at~\ref{subsec:training} we present the training (fitting)
part, at~\ref{subsec:inference} the inference part and
at~\ref{subsec:evaluation} the evaluation part. Finally,
at~\ref{subsec:extensibility} we describe how a user may extend ROMC
with its custom modules.

@@

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}

In this section, we illustrate the implementation performing two
experiments; In chapter~\ref{subsec:ma2} we perform the inference at
the second-order moving average (MA2) model, which is a standard model
of \pkg{ELFI}, and in chapter~\ref{subsec:exec_time} we measure the
execution speed-up achieved by the parallelization of ROMC.

\subsection{Moving average example}
\label{subsec:ma2}
\input{latex_files/ch4/4_2_ma2.tex}

\subsection{Execution time}
\label{subsec:exec_time}
\input{latex_files/ch4/4_3_exec.tex}
